---
title: "Cleaning raw coded data set"
author: "Jan Pfänder & Sacha Altay"
date: "2023-04-27"
output: 
  html_document: 
    keep_md: yes
bibliography: bibliography.bib
---

```{r, message=FALSE}
library(tidyverse)
library(sf) # for map data
```

For our analysis, some of the categories we coded are too fine grained (e.g. there is only one study that used an `accuracy_scale` of 100). Further, we sometimes were not consistent in our coding (e.g. the `country` of some studies was sometimes coded `US`, other times `United States`). In this script, we create new grouped versions of certain variables and code coherent values where needed.

```{r, message=FALSE}
d <- read_csv("./data/data.csv")
```

### Define levels for new moderator variables

```{r}
# define news family groups
political <- c("identity_related", "political" , "political_anti_gov", 
               "political_pro_gov", "politically_balanced", "politically_concordant", 
               "politically_discordant", "politically_neutral", "politically_pro-left", 
               "politically_pro-right", "pro vs anti government")
covid <- c("covid")
other <- c("anything", "environment_climate",  "general", "health", "historical", 
           "medical", "military", "economic", "mixed_health_sex_trump_nature_syria_crime",
           "science")

# define news_format groups
title <- c("title", "title_read_out", "pitch", "title_pitch")
title_picture <- c("title_picture", "title_picture (from real SM)", "title_picture + AUDIO")
title_picture_pitch <- c("title_picture_pitch", "title_picture_pitch_fulltext")
```

## Recode variables and add some new ones

```{r}
# recode variables
d <- d %>% 
  # do some renaming to be in line with variable names used in the 
  # simulated data and thus in the pre-registration
  rename(mean_accuracy_fake = mean_fake_as_accurate, 
         mean_accuracy_true = mean_true_as_accurate, 
         sd_accuracy_fake = SD_fake_as_accurate, 
         sd_accuracy_true = SD_true_as_accurate, 
         n_subj = n
  ) %>% 
  mutate(
    # add a unique identifier variable using `ref` and `sample_id`/`news_id`
    ref = str_replace(ref, " ", "_"),
    unique_sample_id = paste(ref, sampleID, sep = "_"),
    unique_news_id = paste(ref, newsID, sep = "_"),
    # add a numeric version of the accuracy scale variable
    accuracy_scale_numeric = as.numeric(accuracy_scale),
    # add `n_observations` variable
    n_observations = n_subj*n_news,
    # country
    country = ifelse(country == "US", "United States", country),
    country_grouped = ifelse(country == "United States" | country == "Europe/United States", 
                            "US", "non-US"),
    # news_family
    political_concordance = case_when(news_family == "politically_concordant" ~ "concordant", 
                                      news_family == "politically_discordant" ~ "discordant",
                                      TRUE ~ NA_character_), 
    news_family = tolower(news_family),
    news_family_grouped = case_when(news_family %in% all_of(political) ~ "political", 
                                    news_family %in% all_of(covid) ~ "covid",
                                    news_family %in% all_of(other) ~ "other", 
                                    TRUE ~ NA_character_),
    # news_format
    news_format_grouped = case_when(news_format %in% all_of(title) ~ "headline", 
                                    news_format %in% all_of(title_picture) ~ "headline_picture",
                                    news_format %in% all_of(title_picture_pitch) ~ "headline_picture_lede", 
                                    TRUE ~ NA_character_),
    # news_source
    news_source = case_when(news_source == "source" ~ TRUE, 
                            news_source %in% c("no_source", "nosource") ~ FALSE,
                            TRUE ~ NA), 
    # accuracy scale
    accuracy_scale_grouped = ifelse(accuracy_scale %in% c("4", "binary", "6", "7"), 
                                    accuracy_scale, "other"), 
    # perfect symetry
    perfect_symetry = case_when(perfect_symetry == 1 ~ TRUE,
                                perfect_symetry == 0 ~ FALSE,
                                TRUE ~ NA),
    # selection fake news
    selection_fake_news_grouped = ifelse(selection_fake_news == "fact_checking", "fact check sites", "identified by researchers"),
    # for multilevel models, later , we want to identify all observations
    # (i.e. the individual effect sizes) of our data 
    observation_id = 1:nrow(.)
  )
```

### Add map compatible names

We make our country variable compatible to match map data. This involves a couple of coding decisions:

-   One study says "Europe/ United States" - since we don't have more precision on where in Europe, we'll code United States.
-   Hong Kong is not listed as a sovereign state - we'll therefore code as China.

```{r}
# Download "Admin 0 – Countries" from
# https://www.naturalearthdata.com/downloads/110m-cultural-vectors/
world_map <- read_sf("data/ne_110m_admin_0_countries/ne_110m_admin_0_countries.shp") %>%
  # remove Antarctica
  filter(ISO_A3 != "ATA")

# Check which of our country names are not matched by the map data
country_names <- world_map$ADMIN
d %>% reframe(country = unique(country)) %>% filter(!country %in% country_names) 

# # search manually for the country names
# # this was achieved using tab completion like this:
# country_names <- world_map %>% split(.$ADMIN)
# country_names$`United Kingdom`

# make new, map compatible country variable  
d <- d %>% 
  mutate(country_mapcompatible = case_when(country == "United States" ~ "United States of America",
                                           country == "Hong Kong" ~ "China", 
                                           country == "UK" ~ "United Kingdom",
                                           country == "Europe/United States" ~ "United States of America",
                                           TRUE ~ country
                                           )
         )

# add continent variable
d <- left_join(d, world_map %>% select(CONTINENT, ADMIN), by = c("country_mapcompatible" = "ADMIN")) %>% 
  rename(continent = CONTINENT)
```

### Calculate Error

```{r}
# calculate `error` variable
d <- d %>% 
  mutate(
    # calculate error for true news
    error_true = ifelse(accuracy_scale == "binary", 
                        # for binary scales, the maximum is 1
                        1 - mean_accuracy_true, 
                        # for continuous scales, the maximum is the value of
                        # `accuracy_scale_numeric`
                        accuracy_scale_numeric - mean_accuracy_true
                        ),
    # calculate error for fake news
    error_fake = ifelse(
      # For binary scales, the minimum is 0.
      # For standardized continuous scales that reach from 0 to 1, 
      # the minimum is also 0,
      accuracy_scale == "binary" | accuracy_scale == "1", 
      mean_accuracy_fake - 0,
      # for all other continuous scales, the minimum is 1
      mean_accuracy_fake - 1)
  )
```

## Remove Problematic studies

There are a few studies that we collected data for first, but that on second inspection do not meet our inclusion criteria.

```{r}
# # check out papers that have no (neutral) control condition
# d %>% group_by(paperID, ref) %>% 
#   summarize(n = sum(condition == "control")) %>% filter(n < 1) %>% 
#   select(ref)
```

Those studies are:

-   @hoesProminentMisinformationInterventions2023 (made up fake news)

-   @calvilloInitialAccuracyFocus2020 (no neutral control)

-   @baptistaInfluencePoliticalIdeology2021 (they use a very asymmetric scale, and results strongly favour hypotheses H2, one sample finding an effect size as large as d = 2.54 - including it seems like cheating; "According to your knowledge, how do you rate the following headline? on a 5point scale (1---not credible; 2---somehow credible; 3---quite credible; 4---credible; 5---very credible).")

```{r}
d <- d %>% 
  # filter out problematic studies
  filter(!ref %in% c("Hoes_2023", "Calvillo_2020", "Baptista_2021"))
```

## Control condition only

We had collected data for treatment conditions, too. However, this data is irrelevant for this project, so we filter it out here. 

```{r}
# load data
d <- d %>% 
  # control conditions only
  filter(condition == "control")
```


## Write out cleaned data

```{r}
# Export data frame to CSV file with "NA" as NA representation
write_csv(d, "data/cleaned.csv", na = "NA")
```

## References
