---
title: "Revisions: New search string for systematic search"
output:
  pdf_document: default
  html_document:
    keep_md: true
---

```{r, message=FALSE}
library(tidyverse)
library(readxl)
```

# Titles screening

We first want to remove all studies we have looked at in our previous search. 

Load data sheets.
```{r, message=FALSE}
# previous literature search
google <- read_csv("google.csv") 
scopus <- read_csv("scopus.csv") 

# Literature
included <- read_csv("../data/cleaned_before_revisions.csv") %>% 
  # control conditions only
  filter(condition == "control") %>% 
  group_by(paperID) %>% 
  summarize(reference = unique(reference)) %>% 
  arrange(paperID) %>% 
  ungroup() 
excluded <- read_csv("excluded_before_revisions.csv")

# revisions literature search
revisions_google <- read_csv("revisions_scholar.csv", col_types = cols(.default = col_character())) 
revisions_scopus <- read_csv("revisions_scopus.csv", col_types = cols(.default = col_character()))

options(width = 10000)

revisions_scopus  %>% slice(1) %>% select(Abstract) %>% print(width = Inf) 

revisions_scopus  %>% slice(1) %>% select(Abstract) %>% pull
```

Modify string format.
```{r}
# little helper function
make_strings_more_compatible <- function (string) {
  string <- tolower(string)   # no more capital letters
  string <- trimws(string)    # no more white spaces at the end
}

# edit strings
google <- google %>% 
  mutate(across(Title, make_strings_more_compatible))

scopus <- scopus %>% 
  mutate(across(Title, make_strings_more_compatible))

included <- included %>% 
  mutate(across(reference, make_strings_more_compatible))

excluded <- excluded %>% 
  mutate(across(reference, make_strings_more_compatible))

revisions_google <- revisions_google %>% 
  mutate(across(Title, make_strings_more_compatible))

revisions_scopus <- revisions_scopus %>% 
  mutate(across(Title, make_strings_more_compatible))

```

Identify duplicates between revisions and prior search 

```{r}
# extract prior search results
prior_search_doi <- c(scopus$DOI, google$DOI )
prior_search_titles <- c(scopus$Title, google$Title)

# identify duplicates from literature searches
revisions_google <- revisions_google %>% 
  mutate(duplicate_by_DOI = ifelse((DOI %in% prior_search_doi) & !is.na(DOI), TRUE, FALSE),
         duplicate_by_Title = ifelse((Title %in% prior_search_titles), TRUE, FALSE), 
         duplicate_by_either = ifelse((duplicate_by_DOI | duplicate_by_Title) == TRUE, TRUE, FALSE)
         )

# Overview of different matching methods
revisions_google %>% summarise(duplicate_by_DOI = sum(duplicate_by_DOI), 
                     duplicate_by_Title = sum(duplicate_by_Title), 
                     duplicate_by_either = sum(duplicate_by_either))

# store the duplicates with prior search
duplicates_prior_google <- sum(revisions_google$duplicate_by_either)

```

```{r}
# identify duplicates from literature searches
revisions_scopus <- revisions_scopus %>% 
  mutate(duplicate_by_DOI = ifelse((DOI %in% prior_search_doi) & !is.na(DOI), TRUE, FALSE),
         duplicate_by_Title = ifelse((Title %in% prior_search_titles), TRUE, FALSE), 
         duplicate_by_either = ifelse((duplicate_by_DOI | duplicate_by_Title) == TRUE, TRUE, FALSE)
         )

# Overview of different matching methods
revisions_scopus %>% summarise(duplicate_by_DOI = sum(duplicate_by_DOI), 
                     duplicate_by_Title = sum(duplicate_by_Title), 
                     duplicate_by_either = sum(duplicate_by_either))

# store the duplicates with prior search
duplicates_prior_scopus <- sum(revisions_scopus$duplicate_by_either)
```

Add duplicates between the two revisions literature searches 
```{r}
# identify duplicates from literature searches
revisions_google <- revisions_google %>% 
  mutate(duplicate_by_DOI = ifelse((DOI %in% revisions_scopus$DOI) & !is.na(DOI), TRUE, duplicate_by_DOI),
         duplicate_by_Title = ifelse((Title %in% revisions_scopus$Title), TRUE, duplicate_by_Title), 
         duplicate_by_either = ifelse((duplicate_by_DOI | duplicate_by_Title) == TRUE, TRUE, duplicate_by_either)
         )

# Overview of different matching methods
revisions_google %>% summarise(duplicate_by_DOI = sum(duplicate_by_DOI), 
                     duplicate_by_Title = sum(duplicate_by_Title), 
                     duplicate_by_either = sum(duplicate_by_either))

# store duplicates between searches (after removing) prior search results
duplicates_revision <- sum(revisions_google$duplicate_by_either) - duplicates_prior_google

```

We remove all duplicates and store the remaining results in a combined data frame.

```{r}
# make a common search data frame 
revisions_search <- bind_rows(revisions_google %>% 
                                select(DOI, Title, Abstract, duplicate_by_either) %>% 
                                rename(duplicate = duplicate_by_either), 
                              revisions_scopus %>%
                                select (DOI, Title, Abstract, duplicate_by_either)  %>% 
                                rename(duplicate = duplicate_by_either)
                              )

# remove duplicates
revisions_search <- revisions_search %>% 
  filter(duplicate == FALSE) %>% 
  select(-duplicate)

# make title screening data
titles_screening <- revisions_search %>% 
  # make a column for coder decisions
  mutate(include = "")
```

Our search for revisions yielded `r nrow(revisions_google)` entries for the google scholar search and `r nrow(revisions_scopus)` entries for the scopus search. After removing duplicates with prior search results (`r duplicates_prior_google` in the google search, `r duplicates_prior_scopus` in the scopus search), and removing `r duplicates_revision` remaining duplicates between the two searches, we are left with `r nrow(revisions_search)` new titles to screen.

## Export Title screening data

```{r}
write_csv(titles_screening, "revisions_titles_screening.csv")
```

# Abstract screening

We read in the two independently conducted title screening spreadsheets.

```{r}
titles_sacha <- read_csv("revisions_titles_screening_sacha.csv") %>% 
  mutate(across(include, make_strings_more_compatible)) %>% 
  rename(include_sacha = include)

titles_jan <- read_csv("revisions_titles_screening_jan.csv") %>% 
  mutate(across(include, make_strings_more_compatible)) %>% 
  rename(include_jan = include)
```
We then merge this to the originial data, this time adding abstracts. 

```{r}
revisions_search <- bind_cols(revisions_search, 
                              titles_sacha %>% 
                                select(include_sacha), 
                              titles_jan %>% 
                                select(include_jan))
```

Code (dis)agreements. Include all entries where at least one of the coders suggested abstract screening

```{r}
revisions_search <- revisions_search %>% 
  mutate(agreement = ifelse(include_sacha == include_jan, "agreement", "disagreement"), 
         # code abstract screening
         abstract_screening = ifelse(include_jan == "yes" | include_sacha == "yes", TRUE, FALSE)
         )
```

Inspect. 
```{r}
revisions_search %>%
  summarize(across(starts_with("include"), ~ sum(. == "yes")), 
            abstract_screening = sum(abstract_screening), 
            disagreement = sum(agreement == "disagreement")
            )
```
## Remove already included titles

So far we have removed duplicates with the previous search. But we also have added articles that didn't come up in the previous search to our data base. Before going into the abstract, and then full text assessment phase, we want to exclude potential entries of the current, second search, that might already be in our data base.

We also excluded some studies that we came across, but not in the systematic search. We also want to make sure to remove these, before going into the more detailed phase.

```{r}
# Literature
included <- read_csv("../data/cleaned.csv") %>% 
  # control conditions only
  filter(condition == "control") %>% 
  group_by(paperID) %>% 
  summarize(reference = unique(reference)) %>% 
  arrange(paperID) %>% 
  ungroup() 
excluded <- read_csv("excluded.csv")

# make compatible
included <- included %>% 
  mutate(across(reference, make_strings_more_compatible)) %>% 
  select(paperID, reference) %>% 
  mutate(included = "included")

excluded <- excluded %>% 
  mutate(across(reference, make_strings_more_compatible)) %>% 
  select(screening, extent, reference) %>% 
  mutate(included = "excluded")

# Combine the `included` and `excluded` sheet into a single `studies` data frame. 
studies <- bind_rows(included, excluded)
```

We write a function that allows us to identify overlaps between the literature searches. 

The function below checks whether a reference in the second search has a matching title in previously included or excluded articles. It adds the variable(s) `duplicate_*` (by title and/or doi), which evaluates `TRUE` if reference appears in the current, second search, and `FALSE` otherwise. To be more certain that we don't miss anything, we search for matches via both the `DOI` and the `Title` column. 

```{r}
identify_appearance_in_searches <- function(data, compare_by = "DOI and Title") {
  
  if (str_detect(compare_by, "Title")) {
  
 data$duplicate_by_title <- sapply(data$reference, function(x) any(sapply(revisions_search$Title, 
                                                              grepl, x, 
                                                              fixed = TRUE)))
  }
  
  if (str_detect(compare_by, "DOI")) {
  
 data$duplicate_by_DOI <- sapply(data$reference, function(x) any(sapply(revisions_search$DOI, 
                                                              grepl, x, 
                                                              fixed = TRUE)))
  
  }

  if (str_detect(compare_by, "(?=.*DOI)(?=.*Title).*")) {
  data <- data %>%
    mutate(duplicate_by_any = 
             ifelse(
               (duplicate_by_DOI | duplicate_by_title)  == TRUE,
                                        TRUE,
                                        FALSE)
           )
  }
 
 return(data)
  
}
```

We apply this function to the `studies` data frame that contains all studies we assessed. 

```{r}
studies <- identify_appearance_in_searches(studies)
```
Inspect. 
```{r}
studies %>%
  summarize(across(starts_with("duplicate"), ~ sum(., na.rm=TRUE))
            )
```
Filter out duplicates

```{r}
# filter out duplicates
duplicates_titles <- studies %>% 
  filter(duplicate_by_title == TRUE) %>% 
  select(reference) %>% 
  pull()

duplicates_doi <- studies %>% 
  filter(duplicate_by_DOI == TRUE) %>% 
  select(reference) %>% 
  pull()


identify_duplicates <- function(data) {
  # Apply grepl to each string in the "Title" column to check if it's contained in any string of the external vector
  title_check <- sapply(data$Title, function(x) any(grepl(x, duplicates_titles, fixed = TRUE)))
  
  # Apply grepl to each string in the "DOI" column to check if it's contained in any string of the external vector
  doi_check <- sapply(data$DOI, function(x) any(grepl(x, duplicates_doi, fixed = TRUE)))
  
  # Set "duplicate" to TRUE if either the title or the DOI matches
  data$duplicate <- title_check | doi_check
  
  # Return the modified dataset
  return(data)
}

revisions_search <- identify_duplicates(revisions_search)

# inspect
revisions_search %>% 
  filter(duplicate == TRUE) %>% 
  select(Title)
```

```{r}
# remove duplicates
revisions_search <- revisions_search %>% 
  filter(duplicate == FALSE) %>% 
  select(-c(duplicate, include_sacha, include_jan, agreement))
```

Make abstract screening data
```{r}
# make title screening data
abstract_screening <- revisions_search %>% 
  filter(abstract_screening == TRUE) %>% 
  select(-abstract_screening) %>% 
  # make a column for coder decisions
  mutate(include = "")
```


## Export Title screening data

```{r}
# Export data to CSV
write_csv(abstract_screening, "revisions_abstract_screening.csv")
```


# Full text

We read in the two independently conducted abstract screening spreadsheets.

```{r}
abstracts_sacha <- read_csv("revisions_abstracts_sacha.csv") %>% 
  mutate(across(include, make_strings_more_compatible)) %>% 
  rename(include_sacha = include)

abstracts_jan <- read_csv("revisions_abstracts_jan.csv") %>% 
  mutate(across(include, make_strings_more_compatible)) %>% 
  rename(include_jan = include)
```
We then merge this to the orginial data.  

```{r}
abstract_screening <- bind_cols(abstracts_sacha %>% 
                                select(include_sacha), 
                              abstracts_jan %>% 
                                select(include_jan))
```

Code (dis)agreements. Include all entries where at least one of the coders suggested full text assessment

```{r}
abstract_screening <- abstract_screening %>% 
  mutate(agreement = ifelse(include_sacha == include_jan, "agreement", "disagreement"), 
         # code abstract screening
         full_text = ifelse(include_jan == "yes" | include_sacha == "yes", TRUE, FALSE)
         )
```

Inspect. 
```{r}
abstract_screening %>%
  summarize(across(starts_with("include"), ~ sum(. == "yes")), 
            full_text = sum(full_text), 
            disagreement = sum(agreement == "disagreement")
            )
```

Inspect articles that Sacha wants included and Jan not
```{r}
abstract_screening %>%
  filter(include_jan == "no" & include_sacha == "yes")
```

Make full text assessment data.

```{r}
# make title screening data
full_text <- abstract_screening %>% 
  filter(full_text == TRUE) %>% 
  select(-full_text, -starts_with("include"), -agreement)
```


## Export Title screening data

```{r}
# Export data to CSV
write_csv(full_text, "revisions_full_text_assessment.csv")
```






